### 2022-02-28 paper/localpsf.tex
writing

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in detail later, but the key idea is that
\begin{equation}
\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
\end{equation}
if $x_i$ is close to $x$. This is shown in Figure FIG. We therefore approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$. Here $k$ is a user-defined parameter; we use $k=10$. The weights in the linear combination are determined via radial basis function interpolation.

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method. In this method, one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$. Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems.

@@

>>>>

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in Section SECTION, but the key idea is that
\begin{equation}
\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
\end{equation}
if $x_i$ is close to $x$. This is shown in Figure FIG. We therefore approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$. Here $k$ is a user-defined parameter; we use $k=10$. The weights in the linear combination are determined via radial basis function interpolation.

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method. In this method, one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$. Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion. WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A. These H-matrix methods are fast and scale well to large problems.

@@

\subsection{Approximating kernel entries $\Phi(y,x)$}

We approximate kernel entries $\Phi(y,x)$ by preconditioned radial basis function interpolation of the impulse response values
\begin{equation}
\label{eq:phi_mu_shift32}
	\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right),
\end{equation}
where the points $x_i$ are sample points at which $\phi_{x_i}$ is known. Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding points $x_i$ for which \eqref{eq:phi_mu_shift32} is not defined. That is, we exclude points $x_i$ such that $y - \mu(x) + \mu(x_i) \notin \Omega$. In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points. We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.

Once the $k$ nearest sample points, $\{x_i\}_{i=1}^k$, are found, we approximate $\Phi(y,x)$ using preconditioned radial basis function interpolation of the following form:
\begin{equation}
	\Phi(y,x) \approx \sum_{i=1}^k w_i b\left(\|x-x_i\|\right).
\end{equation}
For radial basis functions, $b_i$, we use polyharmonic splines. That is, 
\begin{equation}
b_i(r) = r^d
\end{equation}
if $d$ is odd, and
\begin{equation}
	b_i(r) = \begin{cases}
		r^d \log r, & r \neq 0 \\
		0 & r = 0
	\end{cases}
\end{equation}
if $d$ is even.
The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
\begin{equation}
	Bw = f,
\end{equation}
where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
\begin{equation*}
	B_{ij} = b\left(\|x_i - x_j\|\right),
\end{equation*}
and the right hand side vector $f \in \mathbb{R}^k$ has entries
\begin{equation}
	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
\end{equation}
%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly performing polyharmonic spline radial basis function interpolation of the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$. Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, $V(x)$ varies over six orders of magnitude.

The point $y - \mu(x) + \mu(x_i)$ typically is not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Although it is advisable to avoid evaluating finite element functions at arbitrary points [CITE], here we are forced to do this. We evaluate $\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
\begin{equation}
\label{eq:ellipsoid_check93}
	(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
\end{equation}
If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, so we use $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)=0$ for the purpose of interpolation. If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$. In this case, we look up the batch index $b$ for the batch of sample points containing $x_i$, and evaluate
\begin{equation}
	\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = \eta_b\left(y - \mu(x) + \mu(x_i)\right).
\end{equation}
To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we recall the mesh cell that $y - \mu(x) + \mu(x_i)$ resides in. This mesh cell was already computed as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$. We then evaluate the relevant finite element basis functions on the cell to evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.

If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, and we can use this fact to improve the interpolation accuracy at negligible additional cost.

Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$.

Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing

In the case where $\Aop$ is symmetric, we have
\begin{equation}
	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
\end{equation}
knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
